---
title: "ESM 244 Lab1"
author: "Anastasia Kunz"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)


```


### Read in SF trees data

```{r}
sf_trees<- read.csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic wrangling reminders

refresh skills for data wrangling and summary statistics using functions in the `dplyer` package

find top 5 highest obsv of trees by legal_status, do wrangling, make a graph
using: count


```{r}
top_five_status <- sf_trees %>% 
  count(legal_status) %>% #count does summarize, n function, and group by in one fucntion
  drop_na(legal_status) %>%  # for any variable, just need to specify can be multiple
  rename(tree_count = n) %>%  #NEW NAME GOES ON THE LEFT
  relocate(tree_count) %>%  #single column will move it to the front; must use new name bc rename came first
  slice_max(tree_count, n = 5) #identify rows with high values and only keeps top ones, there is also a min

```

Make a graph by top five observations by legal obsv

```{r}
ggplot(top_five_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count))+ #change to factor associated with tree count associated with that
  geom_col()+
  labs(
    x = "Legal Status",
    y = "Tree Count"
  )+
  coord_flip()+
  theme_minimal()
```


### More data wrangling
only want to keep obsv (rows) blackwood acacia trees

```{r}
blackwood_acacia <-sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>%  #keep observations wehre a certain string is detected ANYWHERE within that variable
  select(legal_status, date, latitude, longitude)

ggplot(blackwood_acacia, aes(x = longitude, y = latitude))+ 
  geom_point()
```

### Tidy r separate and unite functions

usefull for combining and separating columns

common name and scientific name, split them using the :: in the dataset!

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

example of tidyr unite

combine tree and legal status into same column. probaby wouldnt do this

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool_")
```

### Make some actual maps of blackwood acacia trees in SF

use `st_as_sf()` to convert the lat and long to spatial coordinates
set a coord ref system 

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  
```

